% OpenCP Module Player
%
% Documentation LaTeX source
%
% revision history: (please note changes here)
% -doj990413  Dirk Jagdmann  <doj@cubic.org>
%   -initial release

\chapter{Fileselector}
If \cp\ is started without any command line arguments the fileselector will be
loaded. With this powerful tool you can browse through your modules and set
up playlists to be processed by the player. If you switch to the player the
selected files will be loaded and processed.

Files can be stored inside compressed archives to save space on the hard disk.
Those files are automatically unpacked to a temporary directory before
scanning or loading. If you have many modules you should use this feature, as
storing modules inside archives is totally transparent when using the fileselector.

\section{Main screen}
The fileselector splits into three main windows: directory list, playlist
and module information as shown in figure \ref{fileselectorscreen}.

\begin{figure}
\hfill
\setlength{\unitlength}{0.8cm}
\begin{picture}(13.3,10)
\thicklines
\put(0,0){\framebox(13.3,10){}} % ‰uﬂerer Rahmen
\thinlines
\put(0,0){\makebox(13.3,0.5){quickfind}} 
\put(0,0.5){\framebox(13.3,2.5){module information}} 
\put(0,3){\makebox(10,6){directory list}}
\put(10,3){\line(0,6){6}}
\put(10,3){\makebox(3.3,6){play list}}
\put(0,9){\framebox(13.3,0.5){path}} 
\put(0,9.5){\makebox(13.3,0.5){title bar}}
\end{picture}
\hfill\hbox{}
\caption{\label{fileselectorscreen}the fileselector screen}
\end{figure}

The path window shows the current
path and file mask. If you started \cp\ from the home directory you will get
the following: \texttt{C:\symbol{92}OPENCP\symbol{92}*.*} means
that the current directory is \texttt{OPENCP} on your \texttt{C} harddrive
and all files are shown (\texttt{*.*}). You can edit the path and the mask
by pressing \keys{CTRL}+\keys{Enter}. After editing the path press \keys{Enter}
to change to the appropriate directory. You can edit the file mask to include
only some files.
\begin{quote}
\texttt{C:\symbol{92}*.mod} will change to the root directory of hard disk \texttt{C} and
show all files ending with \texttt{.mod}. The default setting should be the
current directory with a file mask \texttt{*.*} to show all files.
\end{quote}

The most important window is the directory list. Here you can see all files
in the current directory. If the extension is known to \cp\ the file information
will be shown in different colors depending on the file type. Files not known
to \cp\ will be shown in standard grey.

Leftmost the file name provided by the operating system is shown.
The extension \texttt{.???} will specify the file type. The next column shows the
title of the file if the file type includes a title. In the third column the
number of channels is displayed. Finally rightmost the filesize is shown in
bytes. If the module is included in a ZIP archive the \emph{real} file size
is displayed.

Right to the directory list you can see the play list. All files listed in
this window will be played, after you change into the player. The order of
entries in this window determine the order in which files are loaded unless you
have enabled the \emph{random} option.

The window at the bottom is the module information. Many music formats can
store general information which is displayed here. If the file type does not
support those information you can edit the fields inside this window manually
and \cp\ will store the information for you.

Finally at the very bottom is the quick find feature, which lets you easily
find files in the current directory.

\section{Usage of the fileselector}
\label{fileselector}
The directory list shows you all files in the current directory which fit to
the file mask set in the path window and \cp\ can detect. Under the alphabetically sorted files
the directories and drives are shown.\footnote{\texttt{@:} is a special drive
which lets you configure \cp\ without editing the \texttt{cp.ini} file, see page \pageref{specialdrive}.}
Use the
\keys{$\uparrow$} and \keys{$\downarrow$} to browse through the files. If you
press \keys{Enter} the selected file will be loaded and played with the player.
Pressing \keys{Enter} while selecting a directory or drive will switch to the
selected item and the directory will be read. \keys{Pgup}, \keys{Pgdown},
\keys{Home} and \keys{End} will work as expected.

If a module is played and you are in the player \keys{f} will beam you to
the fileselector. You can always leave the
fileselector by pressing \keys{Esc} twice! If no module is playing the program
will exit, while you will get back to the player if a module is played in the
background.

Playlists are shown in the playlist window at the right side of the screen. The
currently selected file is appended to the playlist by pressing \keys{$\rightarrow$}
or \keys{Ins}. \keys{$\leftarrow$} or \keys{Del} will remove it again. You can
insert files multiple times into the playlist by pressing the appropriate
keys more than once. If you have files in the playlist exit the fileselector
by pressing \keys{Esc}! This might seem confusing in the beginning, but you
will notice the logic very soon. In the player you start the next song in the
playlist by pressing \keys{Enter}.

Normally you will start the fileselector from the player by pressing \keys{f}.
The current module will continue playing in the background. After you have
selected a file you have to choices:
\begin{itemize}
\item \keys{Enter} will stop the currently played module and load the selected
one. Then you will get back to the player. Use this key if you want to play the
selected module immiediatly.
\item \keys{Esc} will change to the player. Then you can start the next songs
in the playlist by pressing \keys{Enter}. If you have inserted files into the
playlist use this key to exit the fileselector.
\end{itemize}

All files in the current directory will be inserted into the playlist by
pressing \keys{CTRL}+\keys{$\rightarrow$} or \keys{CTRL}+\keys{Ins}. The
playlist will be deleted by pressing \keys{CTRL}+\keys{$\leftarrow$} or
\keys{CTRL}+\keys{Del}.

Although in the playlist window only the filename is shown, \cp\ stores the
complete path information. So you can insert files into the playlist from
totally different directories and drives. If files are inserted into the
playlist you can change to the playlist window by pressing \keys{Tab}. Inside
the playlist window all keys have full functionality. So you can load the
selected module immiediatly by pressing \keys{Enter} or remove the file from
the list by pressing \keys{$\leftarrow$}. If you are in the playlist window
you can move the currently selected file by pressing \keys{CTRL}+\keys{$\uparrow$}
and \keys{CTRL}+\keys{$\downarrow$}. This will affect the order in which files
are processed. \keys{CTRL}+\{\keys{Pgup},\keys{Pgdown},\keys{Home},\keys{End}\}
work as expected.

If many files are inside a directory selecting a module with the cursor keys can
be annoying, because it takes a long time to browse through the list. If you
know the filename you can start typing it on the keyboard. This enables the
quickfind feature. Characters already typed are shown in the quickfind window.
The current directory is searched for files matching the typed characters.
Often you don't have to type the complete filename, as it can be already determined
by the leading chars. The typed characters must not fit the file exactly as
small errors are neglected.

At the bottom of the screen the fileinformation window is located. If the
file includes any additional information it will be shown at the appropriate
fields. You can edit each entry manually. {\small All module information is read by
the fileselector once if it runs along this module the first time. The data
is stored in three files located in your home directory of \cp\ refered as the \emph{module information cache}. If the
fileselector scans a directory and finds a module already stored in the
module information cache it will use the information found in the cache. This
way you can overwrite the information fields of the module. If you change to
a directory which has not been processed by the fileselector it may take some
time to read all file information out of the files and store them in the
module information cache.}

To switch to the module information window press \keys{SHIFT}+\keys{Tab}. You
can use the cursor keys to select the entries. After pressing \keys{Enter}
the information can be edited. When pressing \keys{Enter} again the changes
are stored in the module information cache. Note: \emph{Do not change the entry
\emph{type} as the file could not be loaded properly when the wrong filetype
is entered! Normally you never have to change this entry, except for old
15 instruments amiga noisetracker modules!}

\section{Advanced usage}
\label{fileselectoradvance}
The appearance and behaviour of the fileselector can be edited in the
\texttt{cp.ini} (page \pageref{cp.ini}) file or by pressing \keys{ALT}+\keys{c}. Changes made to the
\texttt{cp.ini} are permanently, while configuration applied with  \keys{ALT}+\keys{c}
is only valid while \cp\ is running.

Afer pressing \keys{ALT}+\keys{c} you can toggle 13 options with keys
\keys{1}\dojdots\keys{9} and \keys{a}\dojdots\keys{d}. The following list
will explain every option:
\begin{itemize}
\item[1] \emph{screen mode:} you can change the screen mode for the fileselector.
80x25 and 80x50 are standard screen modes and should be available on every
vga card. 80x30, 80x60, 132x25, 132x30, 132x50 and 132x60 are only available
with a proper VESA bios installed.
\item[2] \emph{scramble module list order:} if this options is enabled the
files inside the playlist will be played in random order. Otherwise the
order shown in the fileselector from top to bottom will be used.
\item[3] \emph{remove modules from playlist when played:} normally you will
want this enabled as modules are only played once. If you disable this option
you playlist can be processed foreever.
\item[4] \emph{loop modules:} if the music file ends it will start again. The
next file will be played after pressing \keys{Enter}. If you turn off this
option the playlist will play all modules without any user interaction.
\item[5] \emph{scan module information:} When entering a directory the files
are processed to gather module information which can be shown. If you disable
this option directories will be processed quicker.
\item[6] \emph{scan module information files:} the module information cache
in the home directory of \cp\ will be read if this option is enabled.
\item[7] \emph{scan archive contents:} to save hard disk space you can store
your files inside archives like \texttt{ARJ} or \texttt{ZIP}. If the fileselector finds an
archive it will open it to scan for files.
\item[8] \emph{scan module information in archives:} if modules are found
inside archives they will be decrunched to find any module information. This
option can take several minutes if many modules are stored in archives
\item[9] \emph{save module information to disk:} toggles weather to save
gathered informations in the module information chache.
\item[A] \emph{edit window:} If you don't want the module information window
at the bottom disable this option. The directory and playlist windows will
spawn over the complete screen.
\item[B] \emph{module type colors:} different file types are shown in
different colors on the screen. When watched on monochrome monitors or laptops
you might want to disable this option.
\item[C] \emph{module information display mode:} changes the contents of the
directory window. You can also use \keys{ALT}+\keys{tab} or \keys{ALT}+\keys{i}
inside the fileselector.
\item[D] \emph{put archives:} Show archives, so they can be accessed like directories.
Normally this should be disabled if archives are scanned automatically.
\end{itemize}

The screen size can be changed by pressing \keys{ALT}+\keys{z}. In 132 columns
mode some additional module information can be shown in the directory window.
If the fileselector is busy scanning the current directory for files, you can
interrupt the scanning with \keys{ALT}+\keys{s}.

You can delete a file with
\keys{ALT}+\keys{k}. You will be prompted if you really want to delete this
file. When pressing \keys{y} the file will be deleted. If the file is stored
inside an archive the file will be deleted from the archive.\footnote{If
only one file was stored inside a \texttt{.ZIP} archive pkzip will leave an
empty archive of 22 bytes on your harddisk. See section \ref{cp.ini} on how to
avoid this.} The current file can be moved by \keys{ALT}+\keys{m}. You will
have to type the new path in the path window into which the file will be
moved. If the file is inside an archive it will be extracted. You can specify
an existing \texttt{.ARJ} file and the file will be packed into the archive.\footnote{this
works only with ARJ archives by now.}

The module information shown in the module information window can be saved to
a portable ascii file (see page \pageref{mdz}). You may want this feature if you are a composer of
music and want to trade your music together with already processed module
information files. Start the fileselector and edit the information for the
file. Then switch back to the directory window and press \keys{ALT}+\keys{w}.
The fileselector will save a file with the extension \texttt{.MDZ} and the
filename of the selected file, which stores all module information seen in
the module information window. If a directory is scanned and the fileselector
finds such \texttt{.MDZ} files they will be read and processed. The module
information for all files in the current directory can be saved with
\keys{ALT}+\keys{a}. You have to type the filename manually in the path window
without extension!

You may want to change the entry \emph{type} in the module information window
if you have old amiga modules or a non-standard midi file. Very old Noise- and
SoundTracker modules only had 15 instruments and no file identification. So
the fileselector is not able to detect those files as valid modules and refuses
to play them. You have to insert \texttt{M15} in the \emph{type} entry. If
the module does not differ between tempo and speed and is of the 15 instrument
type insert \texttt{M15t}. Some ProTracker modules do not differ between tempo
and speed too. If you have one of those modules use \texttt{MODt}. A module
player for PC called DMP introduced a feature called panning. To enable this
(non-standard) feature insert \texttt{MODd}. If you want to play midi files
with a second drum track on channel 16 use the \texttt{MIDd} option. Any other
file should be autodetected correctly. {\small If you have renamed a module
to a different extension (say \texttt{hello.mod} to \texttt{hello.s3m}) \cp\
will refuse to play it, because the file type is wrong. You could correct this
by inserting the right file type in the module information as shown above. But
it is recommended to rename the file to the right extension instead of tweaking the
autodetetion of the player.}

The current playlist can be saved into the .PLS format by pressing
\keys{ALT}+\keys{p}. You have to type the filename without extension in the path
window. A standard extension \texttt{.PLS} is appended. The playlist can be
loaded just like any other module from the fileselector or at startup.

\label{specialdrive}
The drive \texttt{@:} is a special device which can be used to change the
hardware configuration without leaving the player. If you access this drive
you will see two subdirectories.

In the \texttt{INPUTS} subdirectory you can choose the device which will be
used when sampling from external sources (when playing CD audio tracks or
starting \cp\ in sample mode). The \texttt{DEVICES} directory displays all
devices which where detected at startup. Normally you might want to change
this if you want to save the next file as a \texttt{.WAV} or \texttt{.MP2}
file to the harddisk.\footnote{See section \ref{diskwriter}.}
If you have a soundcard with hardware mixing support
(Gravis, AWE, EWS) you can enable softwaremixing.

If you are the lucky owner of a soundcard capable of hardware sample
playing\footnote{Gravis Ultrasounds, Sound Blaster AWE series and Terratec EWS
series} you only have limited sample memory. If the sample data of the music
is larger than the available memory \cp\ will try to reduce the sample size by
applying the following steps:
\begin{enumerate}
\item Convert 16bit of 8bit samples. This is indicated by a ``!'' in the bit
entry of the instrument section.
\item Half the size of an 8bit sample. This is shown by a small $\frac{1}{2}$.
\item Quarter the size of an 8bit sample, indicated by a $\frac{1}{4}$.
\end{enumerate}

Before \emph{downsampling} the sample data \cp\ will search for the sample with
the lowest frequency spectrum. This sample will be converted first. If enough
space is freed up \cp\ will stop downsampling. If not the loader will continue
until the sample data fits into the sound card memory.

\label{bigmodules}
The above behaviour can be avoided if a file is marked \emph{big} by pressing
\keys{ALT}+\keys{b} in the fileselector. The filesize will turn red. Now a file
will not be loaded into the soundcard memory, but played with the internal
mixing routines. This limits the size of files only to the size of physical
memory.\footnote{You have to apply a valid \emph{playback-} and \emph{mix-device}
to use this feature. See section Configuring on page \pageref{cp.ini} for
details.}

\clearpage
\section{Reference}
\label{fsreference}
\begin{longtable}{r@{ -- }l}
\keys{a}\dojdots\keys{z} & quickfind \\
\keys{ALT}+\keys{a} & write module information \texttt{.mdz} for directory \\
\keys{ALT}+\keys{b} & mark module ``big'' \\
\keys{ALT}+\keys{c} & configure fileselector \\
\keys{ALT}+\keys{d} & goto DOS \\
\keys{ALT}+\keys{i} & change display mode for directory window \\
\keys{ALT}+\keys{k} & delete file \\
\keys{ALT}+\keys{m} & move file \\
\keys{ALT}+\keys{s} & stop scanning module information \\
\keys{ALT}+\keys{w} & write module information \texttt{.mdz} for selected file \\
\keys{ALT}+\keys{z} & toggle screen mode \\
\keys{$\uparrow$}, \keys{$\downarrow$} &  move cursor one entry up/down \\
\keys{CTRL}+\{\keys{$\uparrow$}, \keys{$\downarrow$}\} &  move module up/down on playlist \\
\keys{$\rightarrow$}, \keys{Ins} &  add file to playlist \\
\keys{$\leftarrow$}, \keys{Del} &  remove file from playlist \\
\keys{CTRL}+\{\keys{$\rightarrow$}, \keys{Ins}\} &  add all files to playlist \\
\keys{CTRL}+\{\keys{$\leftarrow$}, \keys{Del}\} &  clear playlist \\
\keys{Pgup}, \keys{Pgdown} &  move cursor one page up/down \\
\keys{CTRL}+\{\keys{Pgup}, \keys{Pgdown}\} &  move module one page up/down in playlist \\
\keys{Home}, \keys{End} &  move cursor to top/bottom of the list \\
\keys{CTRL}+\{\keys{Home}, \keys{End}\} &  move module to top/bottom of playlist \\
\keys{Enter} & play selected file \\
             & change to directory/archive/drive \\
             & edit entry (in module info window) \\
\keys{CTRL}+\keys{Enter} & edit path window \\
\keys{Tab} & change between directory and playlist \\
\keys{ALT}+\keys{Tab} & same as \keys{ALT}+\keys{i} \\
\keys{SHIFT}+\keys{Tab} & change to module info window \\
\keys{Esc} & exit fileselector \\
\end{longtable}

Supported filetypes -- valid options for the \emph{type} entry in the module information window.
\begin{dojlist}
\item[669] 669 Composer module  
\item[AMS] Velvet Studio module  
\item[BPA] Death Ralley archive  
\item[CDA] compact disk CD audio track  
\item[DMF] X Tracker module  
\item[IT] Impulse Tracker module  
\item[MDL] Digi Tracker module  
\item[MID] standard midi file  
\item[MIDd] standard midi file, channel 16 is a second drum track  
\item[MOD] amiga ProTracker 1.1b module  
\item[MODt] amiga ProTracker 1.1b module, effect Fxx is tempo  
\item[MODd] amiga ProTracker 1.1b module with effect 8xx is panning  
\item[MODf] pc Fast Tracker II .mod file  
\item[M15] amiga NoiseTracker module with 15 instruments {\small(plays like ProTracker 1.1b)} 
\item[M15t] amiga NoiseTracker module with 15 instruments, effect Fxx is tempo {\small(plays like ProTracker 1.1b)} 
\item[MP3] MPEG audio format level 1-3  
\item[MTM] Multi Tracker module  
\item[MXM] Mxmplay module  
\item[OKT] Oktalyzer module  
\item[PLS] \cp\ playlist, works also with M3U and PLT playlist files  
\item[PTM] Poly Tracker module  
\item[S3M] Sream Tracker 3 module  
\item[SID] PSID sid file  
\item[UMX] Unreal module file  
\item[ULT] Ultra Tracker module  
\item[WAV] Microsoft RIFF wave file  
\item[WOW] WOW Tracker module  
\item[XM] Fast Tracker 2 module  
\end{dojlist}
